---
title: "Ngono Cost of Resistance"
output: rmarkdown::html_document
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = T, comment = "#>")
options(tibble.print_min = 4, tibble.print_max = 4)
```

```{r}
set.seed(1) 
library(splines)
library(ggplot2)
library(ggtree)
library(reshape2)
library(ResistPhy)
library(ape)
library(posterior)
library(cmdstanr)
library(patchwork)
library(viridis)
library(RColorBrewer)
library(latex2exp)

run_mcmc <- F
```
## Load tree from Yonathan Grad's paper, do some boring bioinformatics stuff
Honestly wish my computer could just read my mind and do what I want, automatically
```{r}
tr <- read.tree(system.file("extdata", "grad2016.nwk", package="ResistPhy", mustWork=T))
tr <- makeNodeLabel(tr)
meta <- read.table(system.file("extdata", "grad2016.tab", package="ResistPhy", mustWork=T),
            sep='\t',comment.char='',header=T,as.is=T)
rownames(meta) <- meta$ID
cutoffs <- read.csv(system.file("extdata", "cutoffs.csv", package="ResistPhy", mustWork=T))
```
Dichotomise Resistance data
```{r}
rdf <- data.frame(ID=meta$ID)
rownames(rdf) <- rownames(meta)

abx_names <- data.frame(cutoff_name = c("CFX", "CIP", "CRO", "AZI", "PEN", "TET", "SPC"),
    meta_name=c("CFX", "CIP", "CRO", "AZI", "PEN", "TET", "SPC")
)

for (s in c(1:nrow(abx_names))) {
    rdf[[abx_names$cutoff_name[s]]] <- (meta[[abx_names$meta_name[s]]] > cutoffs[[abx_names$cutoff_name[s]]][2])
}

rdf<-within(rdf, rm(ID))
```

Select the three lineages of interest
```{r, out.width="100%", dpi=800, fig.align="center"}
CL_1 <- extract.clade(tr, "Node129")$tip.label
CL_2 <- extract.clade(tr, "Node580")$tip.label
CL_3 <- extract.clade(tr, "Node1038")$tip.label

tr_r <- keep.tip(tr, c(CL_1,CL_2,CL_3))

#Remove CFX resistant subclade
tr_r <- drop.tip(tr_r, extract.clade(tr_r, "Node732")$tip.label)

tr_r <- ladderize(tr_r)

p <- ggtree(tr_r) +  geom_nodelab(geom='text',size=4) +
    theme_tree2()
gheatmap(p, rdf, offset=8, width=0.6, 
        colnames=FALSE, legend_title="Resistance")+
        scale_x_ggtree()
```

We are interested in the two lineages within the subtree with root at Node125
```{r, eval=T, out.width="70%", dpi=300, fig.align="center"}
cl_dat <- data.frame(node=nodeid(tr_r, c("Node580", "Node129", "Node1038")), name=c("FQ Susceptible", "FQ Resistant 1","FQ Resistant 2"))
mrsd_full <- max(meta[tr_r$tip.label, "Year"])

p <- ggtree(tr_r, mrsd=paste0(mrsd_full,"-01-01")) + 
    geom_cladelab(
        data = cl_dat,
        mapping = aes(
            node = node, 
            label = name, 
            color = name
        ),
        fontsize = 3,
        angle=90,
        align = TRUE,
        vjust = 0.0,
        hjust = "center",
        offset= 10,
        lwd=4.0,
        show.legend = FALSE
    ) + 
    scale_color_brewer(palette="Dark2") +
    theme_tree2()
h1 <- gheatmap(p, rdf[,c("CIP", "AZI", "PEN", "TET")], width=0.2, colnames=FALSE, legend_title="Resistance")+
        scale_fill_manual(values=c("blue", "red")) + 
        guides(color = "none", fill=guide_legend("MIC >= Cutoff")) +
        scale_x_ggtree() +
        theme(legend.position="bottom",
        axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))

h1
pdf("Figures/gono_tree_90.pdf",8,8)
plot(h1)
dev.off()
```
Node580 corresponds to FQ susceptible, Node129 is FQ resistant
Extract the two lineages, filter out tips with missing FQ profiles,or that are FQ resistant and in the susceptible clade and vice versa
Further remove all tips that are CIP resistant (mostly a small subclade in the susceptible tree)
```{r, eval=T, out.width="100%", dpi=300, fig.align="center"}
sus_tr <- extract.clade(tr_r, "Node580")
res_tr_1 <- extract.clade(tr_r, "Node129")
res_tr_2 <- extract.clade(tr_r, "Node1038")

rdf_sus <- rdf[sus_tr$tip.label,]
rdf_res_1 <- rdf[res_tr_1$tip.label,]
rdf_res_2 <- rdf[res_tr_2$tip.label,]

tip_times_sus <- meta[rownames(rdf_sus),]$Year
tip_times_res_1 <- meta[rownames(rdf_res_1),]$Year
tip_times_res_2 <- meta[rownames(rdf_res_2),]$Year

mrsd_sus <- max(tip_times_sus)
mrsd_res_1 <- max(tip_times_res_1)
mrsd_res_2 <- max(tip_times_res_2)
```
Plot the filtered subtrees
```{r, fig.show="hold", out.width="33%", eval=T}
p1 <- ggtree(sus_tr, mrsd=paste0(mrsd_sus,"-01-01")) +
    geom_tiplab(size=1, align=TRUE, linesize=.5)+
    scale_x_ggtree() + 
    theme_tree2()
gheatmap(p1, rdf, offset=8, width=0.6, 
    colnames=FALSE, legend_title="Resistance")+
    scale_fill_manual(values=c("blue", "red"))+
    labs(title="Susceptible Tree") +
    guides(color = "none", fill=guide_legend("MIC >= Cutoff")) +
    scale_x_ggtree()

p2 <-ggtree(res_tr_1, mrsd=paste0(mrsd_res_1,"-01-01"))+theme_tree2() +
    geom_tiplab(size=1, align=TRUE, linesize=.5)+
    scale_x_ggtree() + 
    theme_tree2()
gheatmap(p2, rdf, offset=8, width=0.6, 
        colnames=FALSE, legend_title="Resistance")+
        scale_fill_manual(values=c("blue", "red"))+
        labs(title="Resistant Tree") +
        guides(color = "none", fill=guide_legend("MIC >= Cutoff")) +
        scale_x_ggtree()

p2 <-ggtree(res_tr_2, mrsd=paste0(mrsd_res_2,"-01-01"))+theme_tree2() +
    geom_tiplab(size=1, align=TRUE, linesize=.5)+
    scale_x_ggtree() + 
    theme_tree2()
gheatmap(p2, rdf, offset=8, width=0.6, 
        colnames=FALSE, legend_title="Resistance")+
        scale_fill_manual(values=c("blue", "red"))+
        labs(title="Resistant Tree") +
        guides(color = "none", fill=guide_legend("MIC >= Cutoff")) +
        scale_x_ggtree()
```

Set the analysis end date, move tips into the middle of the year as opposed to the beginning by offsetting most recent sampling date by +0.5
```{r, eval=T}
t_end <- max(mrsd_res_1, mrsd_res_2, mrsd_sus)+1 
mr_sus <- t_end-mrsd_sus-0.5
mr_res_1 <- t_end-mrsd_res_1-0.5
mr_res_2 <- t_end-mrsd_res_2-0.5

print(t_end)
```
Load and preprocess usage data
```{r}
pl_begin <- 1988
# AZI, PEN, TET, CEF, FQ
dat <- system.file("extdata", "usage-clean.csv", package = "ResistPhy", mustWork = TRUE)
abx_raw <- as.matrix(read.csv(dat, header=F))
abx_df <- as.data.frame(t(abx_raw))
colnames(abx_df) <- c("year",abx_df[1,2:14])
abx_df <- abx_df[-1,]
rownames(abx_df) <- as.integer(abx_df[,1])
abx_df$year <- as.integer(abx_df$year)
abx_df$FQ <- as.numeric(abx_df[, "Ciprofloxacin"]) + as.numeric(abx_df[, "Ofloxacin"])
abx_df$CPH <- as.numeric(abx_df[, "Cefixime"]) + as.numeric(abx_df[, "Ceftriaxone 250 mg"]) + as.numeric(abx_df[, "Ceftriaxone 125 mg"]) +as.numeric(abx_df[, "Other Cephalo."]) 

abx_df$Other <- 100-as.numeric(abx_df[, "FQ"]) - 
                as.numeric(abx_df[, "CPH"]) -
                as.numeric(abx_df[, "Penicillins"]) -
                as.numeric(abx_df[, "Azithromycin 2gm"]) -
                as.numeric(abx_df[, "Tetracyclines"])

usage_f_fq <- yearly_usg_stepfunc(as.numeric(abx_df[sapply(c(pl_begin:t_end), paste0),"FQ"]), c(pl_begin:t_end))
usage_f_pen <- yearly_usg_stepfunc(as.numeric(abx_df[sapply(c(pl_begin:t_end), paste0),"Penicillins"]), c(pl_begin:t_end))
usage_f_tet <- yearly_usg_stepfunc(as.numeric(abx_df[sapply(c(pl_begin:t_end), paste0),"Tetracyclines"]), c(pl_begin:t_end))
usage_f_azi <- yearly_usg_stepfunc(as.numeric(abx_df[sapply(c(pl_begin:t_end), paste0),"Azithromycin 2gm"]), c(pl_begin:t_end))
usage_f_CPH <- yearly_usg_stepfunc(as.numeric(abx_df[sapply(c(pl_begin:t_end), paste0),"CPH"]), c(pl_begin:t_end))
usage_f_other <- yearly_usg_stepfunc(as.numeric(abx_df[sapply(c(pl_begin:t_end), paste0),"Other"]), c(pl_begin:t_end))

time_grid <- seq(from=pl_begin, to=t_end, length.out=4000)
fq_df <- data.frame(year=rep(time_grid,6), 
                    usage=c(sapply(time_grid,usage_f_fq),
                            sapply(time_grid,usage_f_pen),
                            sapply(time_grid,usage_f_tet),
                            sapply(time_grid,usage_f_azi),
                            sapply(time_grid,usage_f_CPH),
                            sapply(time_grid,usage_f_other)),
                    abx=c(rep("Fluoroquinolones",length(time_grid)), 
                          rep("Penicillins", length(time_grid)), 
                          rep("Tetracyclines",length(time_grid)),
                          rep("Azithromycin", length(time_grid)),
                          rep("Cephalosporins", length(time_grid)),
                          rep("Other", length(time_grid))))
```
Plot usage data, plot individual areas for AMRs that the lineages carry resistance against
```{r, out.width="100%", eval=T}
p <- ggplot(fq_df, aes(x=year, y=usage, fill=factor(abx, levels=rev(c("Fluoroquinolones", "Penicillins","Tetracyclines", "Cephalosporins", "Azithromycin", "Other"))))) +
    geom_area() +
    scale_x_continuous(breaks = seq(pl_begin, t_end, by =1)) + 
    labs(x="Year", y="Average Usage as Proportion of Primary Treatment") +
    guides(fill = guide_legend(title="Antimicrobial",reverse=F)) +
    scale_fill_viridis(discrete=T) +
    scale_y_continuous(labels = scales::percent_format(scale = 1), breaks=seq(from=0,to=100, by=10)) +
    geom_vline(xintercept=1995, linetype="longdash", color="gray70",alpha=.8)+
    geom_text(aes(x=1995, label="\nAnalysis Start Date", y=40), size=rel(4.0), colour="gray70", alpha=.8, angle=90)+
    theme_minimal() +
    theme(axis.text.x=element_text(size=rel(0.7), angle = 45, vjust=0.5, hjust=0.5),
        axis.text.y=element_text(size=rel(0.7), hjust=1),
        axis.title.y=element_text(size=rel(0.6)),
        axis.title.x=element_text(size=rel(0.6)),
        aspect.ratio=1,
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.line = element_line(size=rel(0.2), colour = "grey80"),
        legend.position="right",
        legend.justification="left",
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(0,0,0,-10))

pl1 <- h1 + annotation_custom(
    grob = ggplotGrob(p),
    xmin = 1795,
    xmax = 1995,
    ymin = 190,
    ymax = 410)

pdf("Figures/gono_data.pdf",6,6)
pl1
dev.off()
```
We can see that PEN + TET was used in significant quantities before ~1995. The FQ resistant lineage carries resistance against these as well, but susceptible clade does not.
Set the analysis start date to 1995. Subset FQ usage data for the given timespan.
```{r, eval=T}
t_begin <- 1995
usage_df <- data.frame(year=c(t_begin:t_end), usage=as.numeric(abx_df[sapply(c(t_begin:t_end), paste0),"FQ"]))
usage_df$usage <- usage_df$usage/100
```

## Perform Inference
- Time scale is in years
- Infectious period of gono is ~12 days
```{r, eval=T}
gamma_ansatz <- 1/90.0
time_scale <- 365.0
t0 <- t_begin #0
tmax <- t_end #t_end - t_begin
usage_df$time #<- usage_df$time - t_begin
stopifnot(all(usage_df$time<=tmax))
stopifnot(all(usage_df$time>=t0))
```
Sample
```{r, eval=run_mcmc}
out <- infer_costs2(list(sus_tr, res_tr_1, res_tr_2), 
                    list(mr_sus,mr_res_1, mr_res_2),
                    usage_df$usage,
                    usage_df$year,
                    t0,
                    tmax,
                    gamma_ansatz,
                    time_scale,
                    n_iter=2000, 
                    n_warmup=2000,
                    model="nodecay",
                    K=60,
                    L=6.5,
                    seed=1345678,
                    gamma_log_sd = 0.1,
                    stan_control=list(adapt_delta=.99,
                        max_treedepth=13,
                        parallel_chains=4,
                        chains=4,
                        refresh=1000
                    ))
saveRDS(out, "~/gono_90s.rds")
```
Load existing analysis result
```{r, eval=T && !run_mcmc}
out <- readRDS("~/gono_90s.rds")
```
Check convergence
```{r, eval =T}
print(out$converged)
```

### Inferred dynamics & Rt
```{r, eval=T, out.width="100%", dpi=300, fig.align="center"}
source("Figures/figure3.R")
pdf("Figures/figure3_90.pdf",6,6)
plot(fig3)
dev.off()
fig3
```

### Pairs Plot
```{r, eval=T, out.width="100%", dpi=300, fig.align="center"}
fig4 <- plot_qpairs(out)
pdf("Figures/figure4_90.pdf",6,6)
plot(fig4)
dev.off()
fig4
```

### RR map
```{r, eval=T, out.width="100%", dpi=300, fig.align="center"}
source("Figures/figure5.R")
pdf("Figures/figure5_90.pdf",6,6)
plot(fig5)
dev.off()
fig5
```

### Traces & ppcheck
```{r, eval=T, out.width="100%", dpi=300, fig.align="center"}
s4 <- plot_traces(out)
pdf("Figures/s4_90.pdf",12,12)
plot(s4)
dev.off()
s4
```
```{r, eval=T, out.width="100%", dpi=300, fig.align="center"}
s5a <- plot_ppcheck_At(out, 1)
s5b <- plot_ppcheck_At(out, 2)
s5c <- plot_ppcheck_At(out, 3)
pdf("Figures/s5_90.pdf",6,6)
plot(s5a)
plot(s5b)
plot(s5c)
dev.off()
s5a
s5b
s5c
```

### Hyppeparameter pairs
```{r, eval=T, out.width="100%", dpi=300, fig.align="center"}
s6 <- plot_hyperpar_pairs(out)
pdf("Figures/s6_90.pdf",6,6)
plot(s6)
dev.off()
s6
```

```{r, eval=T, out.width="100%", dpi=300, fig.align="center"}
s7a <- plot_epi_dynamic(out,1)
s7b <- plot_rt(out,1)
pdf("Figures/s7_90.pdf",6,6)
plot(s7a)
plot(s7b)
dev.off()
s7a
s7b
```